---
title: "Practical Machine Learning Course Project"
author: "Konstantin Serditov"
date: "14 February 2016"
output: html_document
---

# Problem statement
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. For this project data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants is provided. Participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal is to predict the manner in which they did the exercise.

# Data loading and feature extraction

Setup initial parameters and load libraries.
```{r, message=FALSE}
options(warn=-1) #disable warning messages
Sys.setlocale("LC_TIME", "English") #set English locale
setwd("D:/rprog/ml") #set home directory - raw data is already downloaded there
set.seed(4882) #set.seed for reproducibility
library(caret)
library(lubridate)
training <- read.csv("./pml-training.csv", na.strings=c("","NA"))
testing <- read.csv("./pml-testing.csv", na.strings=c("","NA"))
```

Check percent of values with existing NA values.
```{r}
colNAsum <- sapply(training, function(x) sum(is.na(x)))
colNAsum[colNAsum!=0]/nrow(training)
```

If NA exists in column, 98% of values are NA - let's drop these columns from further consideration
```{r}
dropnames <- names(colNAsum[colNAsum!=0])
training <- training[, ! names(training) %in% dropnames, drop = F]
testing <- testing[, ! names(testing) %in% dropnames, drop = F]
```

There are three timestamp variables: raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp. Full day doesn't seem related to outcome, but it may depend on weekday, hour and minute. Exctract them from cvtd_timestamp and drop original variables.
```{r}
training$cvtd_timestamp <- as.POSIXct(training$cvtd_timestamp, format="%d/%m/%Y %H:%M")
training$weekday <- as.factor(weekdays(training$cvtd_timestamp))
training$hour <- as.factor(strftime(training$cvtd_timestamp, format="%H"))
training$minute <- as.factor(strftime(training$cvtd_timestamp, format="%M"))

testing$cvtd_timestamp <- as.POSIXct(testing$cvtd_timestamp, format="%d/%m/%Y %H:%M")
testing$weekday <- as.factor(weekdays(testing$cvtd_timestamp))
testing$hour <- as.factor(strftime(testing$cvtd_timestamp, format="%H"))
testing$minute <- as.factor(strftime(testing$cvtd_timestamp, format="%M"))

dropnames <- c("raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp")
training <- training[, ! names(training) %in% dropnames, drop = F]
testing <- testing[, ! names(testing) %in% dropnames, drop = F]
```

Drop also X variable since it is just row index:
```{r}
dropnames <- c("X")
training <- training[, ! names(training) %in% dropnames, drop = F]
testing <- testing[, ! names(testing) %in% dropnames, drop = F]
```

# Fit model

Use 10-fold cross validation and random forests to create the model.
```{r}
trctrl <- trainControl(method="cv", number=10, verboseIter = FALSE, savePredictions = "final")
model <- train(classe ~ ., data = training, trControl = trctrl, method = "parRF")
```

# Model errors estimate
```{r}
model$finalModel
```
Model shows good accuracy: 99.87%.

# Predict values for test data set
```{r}
predict(model, testing)
```
